Strong lensing selection effects

### 2023-01-30 paper/seleff.tex
minor changes

%
\usepackage{graphicx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{hyperref}
% To add links in your PDF file, use the package "hyperref"
% with options according to your LaTeX or PDFLaTeX drivers.
%
%\usepackage{url}
%\usepackage{fontawesome}

@@

   \author{Alessandro Sonnenfeld\inst{\ref{sjtu},\ref{leiden}}\and
           Shun-Sheng Li\inst{\ref{leiden}}\and
           Giulia Despali\inst{\ref{ita}}\and
           Anowar Shajib\inst{\ref{chicago},\ref{kicp}}\and
           Edward N. Taylor\inst{\ref{swin}}
          }

   \institute{
Department of Astronomy, School of Physics and Astronomy, Shanghai Jiao Tong University, Shanghai 200240, China\\
              \email{sonnenfeld@sjtu.edu.cn}\label{sjtu}
\and
Leiden Observatory, Leiden University, P.O. Box 9513, 2300 RA Leiden, The Netherlands\label{leiden}\and
Institut für Theoretische Astrophysik, Zentrum für Astronomie, Heidelberg Universität, Albert-Ueberle-Str. 2, 69120, Heidelberg, Germany\label{ita}\and
Department of Astronomy \& Astrophysics, University of Chicago,
Chicago, IL 60637, USA\label{chicago}\and
Kavli Institute for Cosmological Physics, University of Chicago,
Chicago, IL 60637, USA\label{kicp}\and
Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn 3122, Australia\label{swin}\and
             }

@@

The Python code used for the simulation and analysis of the lens sample can be found in a dedicated section of a GitHub repository\footnote{\url{https://github.com/astrosonnen/strong_lensing_tools}}.

@@

A very important quantity for describing the strength of a strong lens is the Einstein radius, $\tein$. For an axisymmetric lens, this is the radius corresponding to the solution of \Eref{eq:lenseq} for a source placed at the same angular position as the lens centre ($\beta=0$).
The circle with radius equal to $\tein$ is known as the tangential critical curve. Images that form here have infinite magnification in the tangential direction.
%
It can be shown that $\tein$ satisfies the following condition,
\begin{equation}\label{eq:criteq}
\bar{\kappa}(<\tein) = 1,
\end{equation}
where $\bar{\kappa}(<\theta)$ is the average surface mass density within a radius equal to $\theta$:
\begin{equation}
\bar{\kappa}(<\theta) \equiv \frac{2}{\theta^2}\int_0^{\theta}d\theta' \kappa(\theta')\theta'.
\end{equation}
%Based on \Eref{eq:criteq},
%we generalise the definition of $\tein$ to the non-axisymmetric case as the circularised radius of the isodensity contour that encloses an average surface mass density equal to $\Sigma_{\mathrm{cr}}$.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{ell_pnt_cs.eps}
\caption{
Point source strong lensing cross-section as a function of lens axis ratio.
Solid lines: cross-section based on the lens event definition of \Eref{eq:pointdet}.
Dashed lines: cross-section based on the detection of four images (quad cross-section).
Lines of different colour correspond to sources of different intrinsic magnitude.
%The lens consists of a galaxy at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.19$, $\reff=4.45$~kpc, $\fdm=0.5$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$.
The parameters of the lens density profile are $\fdm=0.5$, $\gammadm=1.5$, $r_s=10\reff$, $\teff=\tein$.
The baryonic and dark matter components have the same ellipticity and direction of the major axis.
\label{fig:ellpoint_cs}
}
\end{figure}

@@

In the case of an extended source, the complexity of the problem is increased due to the addition of a series of features: the source surface brightness distribution, with its radial profile, shape and orientation, and the point spread function (PSF). 
Moreover, as we discussed in section \ref{ssec:lensdefext}, there are different regimes in strong lensing of extended sources, depending on the relative size of the source and the caustics of the lens.
For this reason, we split our analysis in two parts: first we explore the small source size regime, then we consider cases in which the source size is comparable to or bigger than the lens caustics.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{ell_ext_cs.eps}
\caption{
Strong lensing cross-section of an extended source.
Small source size regime.
The cross-section is plotted as a function of the ratio between the half-light radius of the source and the Einstein radius of the lens.
The lens model is fixed to be the same of \Fref{fig:ellpoint_cs}, with axis ratio $q=0.7$. The source is a circular exponential profile.
Each solid line corresponds to a different value of the total unlensed flux of the source. The flux $f$ is expressed in terms of the background noise rms fluctuation measured over an area equal to $\tein^2$.
The vertical dashed lines correspond to the maximum size for which a galaxy with a given flux can be detected in the absence of lensing.
The horizontal dotted line indicates the cross-section for a very bright point source (i.e. the area enclosed within the caustics).
%Solid grey curves are lines of constant effective surface brightness, $f/(2\pi\theta_s^2)$.
\label{fig:extcs}
}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\columnwidth]{largesource_cs.eps}
\caption{
Strong lensing cross-section of an extended source. Large source size regime.
The cross-section is plotted as a function of the ratio between the Einstein radius of the lens and the half-light radius of the source.
The source model is fixed to a circular S\'{e}rsic profile with $n=1$, detected out to the half-light radius.
The source half-light radius measures $20$ pixels, and the PSF has an FWHM of two pixels.
The lens model is varied: starting from the model of \Fref{fig:extcs}, the critical surface mass density is increased to produce progressively smaller values of $\tein$.
Solid line: lensing cross-section, in units of the footprint area of the source in the absence of lensing, as a function of the ratio between the lens Einstein radius and the source half-light radius.
Dashed line: area enclosed within the lens caustics.
Dotted line: value of the cross-section corresponding to the area of the source.
Dash-dotted line: FWHM of the PSF.
\label{fig:largesourcecs}
}
\end{figure}

@@

%We randomly assign a value of $\asps$ to each galaxy in the sample, by drawing it from the following distribution:
For each galaxy in the sample, we randomly drew a value of $\log{\asps}$ from the following distribution:
\begin{equation}
\pr(\log{\asps}) \sim \mathcal{N}(0.1, \sigma_\alpha^2),
\end{equation}
where the notation $\mathcal{N}(\mu,\sigma^2)$ indicates a Gaussian with mean $\mu$ and variance $\sigma^2$.
We set the mean of $\log{\asps}$ to $0.1$, as this is an intermediate value among estimates of $\asps$ from the literature \citep{CvD12, Cap++13, SLC15, Son++15, Son++19}.
We adopted a few different values for the scatter $\sigma_\alpha$. We explain in \ref{ssub:scat} how these were chosen.

@@

Our background galaxy population is taken from the \textsc{surfs}-based KiDS-Legacy-Like Simulation (SKiLLS) input catalogue, a hybrid simulation catalogue integrating cosmological simulation with high-quality imaging observations (Li et al., in prep). The cosmological simulation is obtained from the Synthetic UniveRses For Surveys (\textsc{surfs}) simulations, a set of $N$-body simulations from \citet{Ela++18}. The galaxy properties, including the star formation history and the metallicity history, are from an open-source semi-analytic model named \textsc{Shark}\footnote{\url{https://github.com/ICRAR/shark}}~\citep{Lag++18}. The original photometry is drawn from a stellar population synthesis technique using stellar synthesis libraries with physically motivated dust attenuation and re-emission models~\citep{Rob++20}. Li et al. (in prep) further applied an empirical correction to the original synthetic photometry to better agree with the COSMOS2015 observations~\citep{Lai++16}. The galaxy morphology is described by a S\'ersic profile with three parameters: the half-light radius in angular units $\theta_{
mathrm{e,s}}$, the S\'ersic index $n_\mathrm{s}$, and the axis ratio $q_\mathrm{s}$. These structural parameters are learned from the imaging data obtained with the Advanced Camera for Surveys (ACS) instrument on the \textit{Hubble} Space Telescope~\citep{Gri++12}. We refer to Li et al. (in prep) for details on the learning algorithm and validation.

@@

For each of the three intrinsic scatter scenarios, we drew a population of foreground galaxies covering $1000$ square degrees.
The expectation value of the number of galaxies in the corresponding volume, given the redshift and stellar mass cuts described in section \ref{ssub:mstarz}, is around $300000$.
%Then we generated a population of sources behind the lenses.
%We approximated the foreground galaxies as isolated: for each source, we only modelled the contribution to the lensing field generated by a single lens.
We approximated the foreground galaxies as isolated: when determining whether a galaxy acts as a strong lens, we only modelled the contribution to the lensing signal from the galaxy itself, and neglected that of the environment.
We discuss the possible implications of this approximation in \Sref{sect:discuss}.
For each lens, we determined its caustics relative to the highest source redshift, $\zsource=2.5$. We then placed sources randomly behind the lens. If at least one source fell within a circular region enclosing the caustics, we proceeded to compute its lensed images.

@@

For the simulation with extended sources, we produced images with properties similar to those expected for the Euclid Wide survey 
citep{Sca++22}.
We used a pixel size of $0.1''$ and we applied a Moffat PSF with an FWHM of $0.2''$ and a $\beta$ parameter of $5.0$.
Finally, we assumed a background noise level such that an extended source with half-light radius $0.5''$ and an apparent magnitude in the absence of lensing of $\msource=25$ is detected with $S/N=10$.
%These values are close to those expected for the Euclid Wide survey \citep{Sca++22}.
%
We then applied the peak detection-based lens selection criterion introduced in section \ref{ssec:lensdefext} to find the strong lenses.

@@

In this section we present the results of the lens population simulations. 
Section \ref{ssec:galbias} shows the results of the galaxy-galaxy lens experiment, while section \ref{ssec:quasarbias} focuses on the population of galaxy-quasar lenses.
Given the number of parameters that are needed to describe our model, providing a complete characterisation of the strong lensing bias is a problem with relatively high dimensionality, and is beyond the scope of this paper.
For the sake of conciseness, we focus instead on the quantities that we consider most important.
Nevertheless, the output of our simulations is available online\footnote{\url{https://github.com/astrosonnen/sl_selection_effects}}. We encourage readers who are interested in studying aspects of the strong lensing bias that are not covered in this section to download our data and analyse them directly.

@@

In order for a galaxy-source pair to be included as a strong lens in a survey, three condition must be met.
First of all, at least part of the source must be multiply imaged.
Second, the multiple images must be detectable.
Third, the lens must be recognised as such.
Each one of these conditions introduces a bias with respect to the parent population of foreground galaxies and background sources. Together, they define the lens selection probability term $\psel$ of \Eref{eq:one}.
%In the best case scenario, the only sources of bias are the first two points.
The first two points are intrinsic to a strong lensing survey and constitute an unavoidable source of bias. 
The best case scenario occurs when the efficiency of including a detected strong lens in a survey is always one, and the sample is $100\%$ complete. In this case, the third condition does not introduce any further selection and the strong lensing bias is minimised.
We explored this scenario in \Sref{sect:results} when setting the minimum Einstein radius to zero.
If only lenses with Einstein radius larger than a given threshold are selected, however, the bias generally increases.

@@

As we discussed throughout this paper, the amplitude of the strong lensing bias depends on the intrinsic scatter in the mass parameters of the foreground galaxy population.
One way to minimise the bias, therefore, is to identify scaling relations between observable quantities and mass-related properties that can account for part of the scatter.
Describing the inner dark matter distribution as a function of stellar mass and half-light radius, as we did in section \ref{ssec:galbias}, is the first step in this direction: the bias in $\mdmfive$ at fixed $\mstar$ and $\reff$ is smaller than the overall shift in the median $\mdmfive$ of the lens population (albeit by a marginal amount, as we pointed out earlier).
This description could be extended by including the central velocity dispersion as an additional control parameter. The velocity dispersion is directly related to the mass distribution of the lens galaxy, which means that, for example, the distribution in $\asps$ at fixed velocity dispersion should be narrower than its global distribution marginalised over the whole population. 
The central velocity dispersion, however, is very sensitive to the orbital anisotropy and to gradients in stellar mass-to-light ratio, which are not well known, therefore it is difficult to quantitatively estimate the benefit of including it in the description of the lensing bias.

@@

The simulations on which our analysis is based are as complex as required by the goal of the analysis itself, which is to estimate the amplitude of the strong lensing bias in a few key quantities.
We did, however, make some simplifying assumptions.
One such assumption consisted in neglecting the contribution from line-of-sight structure and the environment to the lensing signal, which typically introduce an external shear and convergence.
The effect of including external shear in a lens model is similar to that of changing the ellipticity of the lens. %Therefore the main result that would be affected by neglecting it is the bias on the axis ratio.
External convergence mimics the effect of adding or removing a constant sheet of invisible mass, producing an effect similar to varying the distribution of dark matter.
Typical values of the external shear and external convergence are $|\gamma|<0.1$ and $|\kappa_{\mathrm{ext}}|<0.1$ \citep{Mil++20}: these values are small compared to the typical ellipticities and dark matter fractions of our simulated lenses.
Therefore, while including them might modify slightly the amplitude of the strong lensing bias in the axis ratio and dark matter distribution parameters, the conclusions our our analysis would not be affected.

@@

The collaboration leading to this work was initiated at the 2022 Lorentz Center workshop “Bridging gaps between dynamical probes of galaxies”.

>>>>

%
\usepackage{graphicx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{hyperref}
\usepackage{siunitx}
% To add links in your PDF file, use the package "hyperref"
% with options according to your LaTeX or PDFLaTeX drivers.
%
%\usepackage{url}
%\usepackage{fontawesome}

@@

   \author{Alessandro Sonnenfeld\inst{\ref{sjtu},\ref{leiden}}\and
           Shun-Sheng Li\inst{\ref{leiden}}\and
           Giulia Despali\inst{\ref{ita}}\and
           Anowar Shajib\inst{\ref{chicago},\ref{kicp},\ref{einstein}}\and
           Edward N. Taylor\inst{\ref{swin}}
          }

   \institute{
Department of Astronomy, School of Physics and Astronomy, Shanghai Jiao Tong University, Shanghai 200240, China\\
              \email{sonnenfeld@sjtu.edu.cn}\label{sjtu}
\and
Leiden Observatory, Leiden University, P.O. Box 9513, 2300 RA Leiden, The Netherlands\label{leiden}\and
Institut für Theoretische Astrophysik, Zentrum für Astronomie, Heidelberg Universität, Albert-Ueberle-Str. 2, 69120, Heidelberg, Germany\label{ita}\and
Department of Astronomy \& Astrophysics, University of Chicago,
Chicago, IL 60637, USA\label{chicago}\and
Kavli Institute for Cosmological Physics, University of Chicago,
Chicago, IL 60637, USA\label{kicp}\and
NHFP Einstein Fellow\label{einstein}\and
Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn 3122, Australia\label{swin}\and
             }

@@

The Python code used for the simulation and analysis of the lens sample can be found in a dedicated section of a GitHub repository\footnote{\url{https://github.com/astrosonnen/strong_lensing_tools/papers/selection_effects}}.

@@

A very important quantity for describing the strength of a strong lens is the Einstein radius, $\tein$. For an axisymmetric lens, this is the radius corresponding to the solution of \Eref{eq:1dlenseq} for a source placed at the same angular position as the lens centre ($\beta=0$).
The circle with radius equal to $\tein$ is known as the tangential critical curve. Images that form here have infinite magnification in the tangential direction.
%
It can be shown that $\tein$ satisfies the following condition,
\begin{equation}\label{eq:criteq}
\bar{\kappa}(<\tein) = 1,
\end{equation}
where $\bar{\kappa}(<\theta)$ is the average surface mass density within a radius equal to $\theta$:
\begin{equation}
\bar{\kappa}(<\theta) \equiv \frac{2}{\theta^2}\int_0^{\theta}d\theta' \kappa(\theta')\theta'.
\end{equation}
%Based on \Eref{eq:criteq},
%we generalise the definition of $\tein$ to the non-axisymmetric case as the circularised radius of the isodensity contour that encloses an average surface mass density equal to $\Sigma_{\mathrm{cr}}$.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{ell_pnt_cs.eps}
\caption{
Point source strong lensing cross-section as a function of lens axis ratio.
Solid lines: cross-section based on the lens event definition of \Eref{eq:pointdet}.
Dashed lines: cross-section based on the detection of four images (quad cross-section).
Lines of different colour correspond to sources of different intrinsic magnitude.
The dashed blue line, corresponding to the brightest source magnitude, overlaps completely with the dashed orange line.
The dashed purple line is zero: very faint sources cannot produce any detectable quads.
%The lens consists of a galaxy at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.19$, $\reff=4.45$~kpc, $\fdm=0.5$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$.
The parameters of the lens density profile are $\fdm=0.5$, $\gammadm=1.5$, $r_s=10\reff$, $\teff=\tein$.
The baryonic and dark matter components have the same ellipticity and direction of the major axis.
\label{fig:ellpoint_cs}
}
\end{figure}

@@

In the case of an extended source, the complexity of the problem is increased due to the addition of a series of features: the source surface brightness distribution, with its radial profile, shape and orientation, and the point spread function (PSF). 
Moreover, as we discussed in section \ref{ssec:lensdefext}, there are different regimes in strong lensing of extended sources, depending on the relative size of the source and the caustics of the lens.
For this reason, we split our analysis into two parts. First, we explore the small source size regime, where the source size is comparable to or smaller than the lens caustics. Then, we consider cases in which the source size is bigger than the lens caustics, which we refer to as the large source regime.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{ell_ext_cs.eps}
\caption{
Strong lensing cross-section of an extended source.
Small source size regime ($\theta_{\mathrm{e,s}} < \tein$).
The cross-section is plotted as a function of the ratio between the half-light radius of the source and the Einstein radius of the lens.
The lens model is fixed to be the same of \Fref{fig:ellpoint_cs}, with axis ratio $q=0.7$. The source is a circular exponential profile.
Each solid line corresponds to a different value of the total unlensed flux of the source. The flux $f$ is expressed in terms of the background noise rms fluctuation measured over an area equal to $\tein^2$.
The vertical dashed lines correspond to the maximum size for which a galaxy with a given flux can be detected in the absence of lensing.
The horizontal dotted line indicates the cross-section for a very bright point source (i.e. the area enclosed within the caustics).
%Solid grey curves are lines of constant effective surface brightness, $f/(2\pi\theta_s^2)$.
\label{fig:extcs}
}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\columnwidth]{largesource_cs.eps}
\caption{
Strong lensing cross-section of an extended source. Large source size regime ($\theta_{\mathrm{e,s}} \gtrsim \tein$).
The cross-section is plotted as a function of the ratio between the Einstein radius of the lens and the half-light radius of the source.
The source model is fixed to a circular S\'{e}rsic profile with $n=1$, detected out to the half-light radius.
The source half-light radius measures $20$ pixels, and the PSF has an FWHM of two pixels.
The lens model is varied: starting from the model of \Fref{fig:extcs}, the critical surface mass density is increased to produce progressively smaller values of $\tein$.
Solid line: lensing cross-section, in units of the footprint area of the source in the absence of lensing, as a function of the ratio between the lens Einstein radius and the source half-light radius.
Dashed line: area enclosed within the lens caustics.
Dotted line: value of the cross-section corresponding to the area of the source.
Dash-dotted line: FWHM of the PSF.
\label{fig:largesourcecs}
}
\end{figure}

@@

%We randomly assign a value of $\asps$ to each galaxy in the sample, by drawing it from the following distribution:
For each galaxy in the sample, we randomly drew a value of $\log{\asps}$ from the following distribution:
\begin{equation}
\pr(\log{\asps}) \sim \mathcal{N}(0.1, \sigma_\alpha^2),
\end{equation}
where the notation $\mathcal{N}(\mu,\sigma^2)$ indicates a Gaussian with mean $\mu$ and variance $\sigma^2$.
We set the mean of $\log{\asps}$ to $0.1$, as this is an intermediate value among estimates of $\asps$ from the literature \citep{CvD12, Cap++13, SLC15, Son++15, Son++19}.
We adopted a few different values for the scatter $\sigma_\alpha$. We explain in Section \ref{ssub:scat} how these were chosen.

@@

Our background galaxy population is taken from the \textsc{surfs}-based KiDS-Legacy-Like Simulation (SKiLLS) input catalogue, a hybrid simulation catalogue integrating cosmological simulation with high-quality imaging observations (Li et al., in prep). The cosmological simulation is obtained from the Synthetic UniveRses For Surveys (\textsc{surfs}) simulations, a set of $N$-body simulations from \citet{Ela++18}. The galaxy properties, including the star formation history and the metallicity history, are from an open-source semi-analytic model named \textsc{Shark}\footnote{\url{https://github.com/ICRAR/shark}}~\citep{Lag++18}. The original photometry is drawn from a stellar population synthesis technique using stellar synthesis libraries with physically motivated dust attenuation and re-emission models~\citep{Rob++20}. Li et al. (in prep) further applied an empirical correction to the original synthetic photometry to better agree with the COSMOS2015 observations~\citep{Lai++16}. The galaxy morphology is described by a S\'ersic profile with three parameters: the half-light radius in angular units $\theta_{
\mathrm{e,s}}$, the S\'ersic index $n_\mathrm{s}$, and the axis ratio $q_\mathrm{s}$. These structural parameters are learned from the imaging data obtained with the Advanced Camera for Surveys (ACS) instrument on the \textit{Hubble} Space Telescope~\citep{Gri++12}. We refer to Li et al. (in prep) for details on the learning algorithm and validation.

@@

For each of the three intrinsic scatter scenarios, we drew a population of foreground galaxies covering $1000$ square degrees.
The expectation value of the number of galaxies in the corresponding volume, given the redshift and stellar mass cuts described in section \ref{ssub:mstarz}, is around \num{300000}.
%Then we generated a population of sources behind the lenses.
%We approximated the foreground galaxies as isolated: for each source, we only modelled the contribution to the lensing field generated by a single lens.
We approximated the foreground galaxies as isolated: when determining whether a galaxy acts as a strong lens, we only modelled the contribution to the lensing signal from the galaxy itself, and neglected that of the environment.
We discuss the possible implications of this approximation in \Sref{sect:discuss}.
For each lens, we determined its caustics relative to the highest source redshift, $\zsource=2.5$. We then placed sources randomly behind the lens. If at least one source fell within a circular region enclosing the caustics, we proceeded to compute its lensed images.

@@

For the simulation with extended sources, we produced images with properties similar to those expected for the Euclid Wide survey \citep{Sca++22}.
We used a pixel size of $0.1''$ and we applied a Moffat PSF with an FWHM of $0.2''$ and a $\beta$ parameter of $5.0$.
Finally, we assumed a background noise level such that an extended source with half-light radius $0.5''$ and an apparent magnitude in the absence of lensing of $\msource=25$ is detected with $S/N=10$.
%These values are close to those expected for the Euclid Wide survey \citep{Sca++22}.
%
We then applied the peak detection-based lens selection criterion introduced in section \ref{ssec:lensdefext} to find the strong lenses.

@@

In this section we present the results of the lens population simulations. 
Section \ref{ssec:galbias} shows the results of the galaxy-galaxy lens experiment, while section \ref{ssec:quasarbias} focuses on the population of galaxy-quasar lenses.
Given the number of parameters that are needed to describe our model, providing a complete characterisation of the strong lensing bias is a problem with relatively high dimensionality, and is beyond the scope of this paper.
For the sake of conciseness, we focus instead on the quantities that we consider most important.
Nevertheless, the output of our simulations is available online\footnote{\url{https://github.com/astrosonnen/strong_lensing_tools/papers/selection_effects}}. We encourage readers who are interested in studying aspects of the strong lensing bias that are not covered in this section to download our data and analyse them directly.

@@

In order for a galaxy-source pair to be included as a strong lens in a survey, three conditions must be met.
First of all, at least part of the source must be multiply imaged.
Second, the multiple images must be detectable.
Third, the lens must be recognised as such.
Each one of these conditions introduces a bias with respect to the parent population of foreground galaxies and background sources. Together, they define the lens selection probability term $\psel$ of \Eref{eq:one}.
%In the best case scenario, the only sources of bias are the first two points.
The first two points are intrinsic to a strong lensing survey and constitute an unavoidable source of bias. 
The best case scenario occurs when the efficiency of including a detected strong lens in a survey is always one, and the sample is $100\%$ complete. In this case, the third condition does not introduce any further selection and the strong lensing bias is minimised.
We explored this scenario in \Sref{sect:results} when setting the minimum Einstein radius to zero.
If only lenses with Einstein radius larger than a given threshold are selected, however, the bias generally increases.

@@

As we discussed throughout this paper, the amplitude of the strong lensing bias depends on the intrinsic scatter in the mass parameters of the foreground galaxy population.
One way to minimise the bias, therefore, is to identify scaling relations between observable quantities and mass-related properties that can account for part of the scatter.
Describing the inner dark matter distribution as a function of stellar mass and half-light radius, as we did in section \ref{ssec:galbias}, is the first step in this direction: the bias in $\mdmfive$ at fixed $\mstar$ and $\reff$ is smaller than the overall shift in the median $\mdmfive$ of the lens population (albeit by a marginal amount, as we pointed out earlier).
This description could be extended by including the central velocity dispersion as an additional control parameter. 
%Galaxies selected in velocity dispersion are similar to existing samples of lenses in terms of other observable quantities
The velocity dispersion is directly related to the mass distribution of the lens galaxy, which means that, for example, the distribution in $\asps$ at fixed velocity dispersion should be narrower than its global distribution marginalised over the whole population. 
The central velocity dispersion, however, is very sensitive to the orbital anisotropy, to the three-dimensional structure, and to gradients in stellar mass-to-light ratio, which are not well known, therefore it is difficult to quantitatively estimate the benefit of including it in the description of the lensing bias.

@@

The simulations on which our analysis is based are as complex as required by the goal of the analysis itself, which is to estimate the amplitude of the strong lensing bias in a few key quantities.
We did, however, make some simplifying assumptions.
One such assumption consisted in neglecting the contribution from line-of-sight structure and the environment to the lensing signal, which typically introduce an external shear and convergence.
The effect of including external shear in a lens model is similar to that of changing the ellipticity of the lens. %Therefore the main result that would be affected by neglecting it is the bias on the axis ratio.
External convergence mimics the effect of adding or removing a constant sheet of invisible mass, producing an effect similar to varying the distribution of dark matter.
Typical values of the external shear and external convergence are $|\gamma|<0.1$ and $|\kappa_{\mathrm{ext}}|<0.1$ \citep{Mil++20}: these values are small compared to the typical ellipticities and dark matter fractions of our simulated lenses.
Therefore, while including them might modify slightly the amplitude of the strong lensing bias in the axis ratio and dark matter distribution parameters, the conclusions of our analysis would not be affected.

@@

The collaboration leading to this work was initiated at the 2022 Lorentz Center workshop “Bridging gaps between dynamical probes of galaxies”.
AS and SSL are supported by NOVA, the Netherlands Research School for Astronomy.
Support for this work was provided by NASA through the NASA Hubble Fellowship grant HST-HF2-51492 awarded to AJS by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., for NASA, under contract NAS5-26555.
